General constraints on magic distillation protocols from statistical mechanics

### 2021-05-04 appendix.tex

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

@@

>>>>

\begin{theorem}
	Consider a magic distillation protocol transforming noisy Strange state
	\begin{equation*}
		\rho_S(\epsilon)^{\otimes n} \longrightarrow \E(\rho_S(\epsilon)^{\otimes n})=\rho_S(\epsilon')^{\otimes m} 
	\end{equation*}
	with $n, m \gg 1$.

@@

\newpage
\subsection{Free energy dependent magic distillation bounds}\label{free-energy-bound-proof}
Here we prove theorem \ref{thm:free-energy}, which bounds magic distillation rates in terms of free energies.

\begin{theorem}
	Consider a magic distillation protocol on qutrits that transforms $n$ copies of an $\epsilon$--noisy Strange state into $m$ copies of an $\epsilon'$--noisy Strange state, with depolarising errors $\epsilon' \leq \epsilon \leq 3/7$.

	Let $T =(k\beta)^{-1}$ be any temperature for the physical system and let $H= \sum_{k \in \mathbb{Z}_3} E_k |E_k\>\<E_k|$ be the Hamiltonian of each qutrit subsystem in its eigen-decomposition.
Assume that for $n,m \rightarrow \infty$ the protocol's channel generates negligible correlations on the equilibrium state $\tau$, and so $\tau^{\otimes n} \longrightarrow \tau^{\prime \otimes m}$ for $n,m \gg 1$. We write the state $\tau'$ as $\tau' = e^{-\beta H'}/\Z'$ for some Hermitian $H'$.

Given this, there are local Clifford changes of basis $C_1,C_2$, such that $\rho_1 := C_1 \rho_S(\epsilon) C_1^\dagger$ and $\rho_2 := C_2 \rho_S(\epsilon') C_2^\dagger$ for which the protocol gives
\begin{equation}
\rho_1^{\otimes n} \longrightarrow \rho_2^{ \otimes m}.
\end{equation}
Moreover, the magic distillation rate $R = m/n$ for the protocol is bounded by the expression
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\phi' - F')},
\end{equation}
where $F$ is the free energy of $\tau$,  and 
\begin{equation}
\phi = -\beta^{-1} \log \zeta
\end{equation}
with $\zeta$ given by the equations
\begin{align}
\zeta &= \sum_{k\in \mathbb{Z}_3} \alpha_k e^{-\beta E_k}, \nonumber\\
\alpha_k &= \sum_{r \in \mathbb{Z}_3} \braket{E_k}{-r}\braket{r}{E_k}.
\end{align}
The primed variables are defined similarly for the output system.
\end{theorem}

\begin{proof}	
For the sake of clarity, we write $\rho_n \coloneqq \rho_S(\epsilon)^{\otimes n}$, $\rho'_m \coloneqq \rho_S(\epsilon')^{\otimes m}$, $\tau_n \coloneqq \tau^{\otimes m}$ and $\sigma_m \coloneqq \sigma^{\otimes m}$.

To establish the distillation bound we consider the distillation protocol that gives $\E( \rho_n) = \rho_m$ for the magic states. 
We then consider that the protocol transforms the reference equilibrium state as $\E( \tau_n) = \sigma_m$.
Since $\tau$ and $\sigma$ are assumed to be full rank stabilizer states, they have a strictly positive Wigner distribution, while, in contrast, the input and output magic states generally have quasi-probability Wigner distributions. 
For any such protocol we therefore have that
\begin{equation}
	( W_{\rho_n}(\bmx), W_{\tau_n}(\bmx) ) \succ ( W_{\rho'_m}(\bmx), W_{\sigma_m}(\bmx) ),
\end{equation}
or, equivalently, in terms of the relevant Lorenz curves,
\begin{equation}
	L_{\rho_n |\tau_n}(x) \ge L_{\rho'_m |\sigma_m}(x) \mbox{ for all } x.
\end{equation}

We define the rescaled Wigner distribution $W_{\rho | \tau}(\x) \coloneqq W_\rho(\x)/W_\tau(\x)$, which is always well-defined since $\tau$ is full-rank. 
Due to the multiplicative property of the Wigner distribution, the rescaled distribution is also multiplicative in the sense that
\begin{equation}
	W_{\rho \otimes \rho' | \tau \otimes \tau'} (\x_1 \oplus \x_2) = W_{\rho | \tau}(\x_1)W_{\rho' | \tau'}(\x_2),
\end{equation}
for any states $\rho, \rho'$ and any full-rank stabilizer states $\tau, \tau'$.
Therefore we have that
\begin{align}
	W_{\rho_n |\tau_n} (\x) &= \prod_{i=1}^n W_{\rho|\tau}(\x_i)\\
	W_{\rho_n } (\x) &= \prod_{i=1}^n W_\rho(\x_i)
\end{align}
where $\x = \oplus_{i=1}^n \x_i \in \mathbb{Z}_3^n$ is the phase space point for the full system in terms of those of the individual subsystems.

The points defining the Lorenz curve $L_{\rho_n |\tau_n}(x)$ are obtained from sorting the components of $W_{\rho_n |\tau_n}(\x)$ in non-increasing order and then computing the partial sums of $W_{\rho_n |\tau_n}(\pi(\x))$ where $\pi$ is the permutation that realises the sorting. 
However, similarly to the unital fragment analysis, we aim to use the constraint that is obtained by considering the line segment connecting the origin to the first elbow of both Lorenz curves.
The proof of this constraint is provided in~\cref{app:elb_constraints}.
Therefore, we focus on finding the coordinates of some point on this line segment of the Lorenz curves.

We find the coordinates for the input Lorenz curve, but the exact same analysis can be applied on the output Lorenz curve.
The Wigner distribution of a single noisy Strange state consists of 1 negative component $W_{\rho}(0,0) = -v(\epsilon)$ and 8 positive components $W_{\rho}(\bmx) = u(\epsilon)$ for $\bmx \neq \bmo$.
The Wigner distribution of the full-rank, stabilizer equilibrium state $\tau$ is $W_{\tau}(\bmx) > 0$ for all $\bmx \in \mathbb{Z}_3^2$.
We keep the components completely general.

Assume that the smallest component of the equilibrium state distribution is $W_{\tau}(i,j)$.
We then make use of the freedom to apply Clifford unitaries in order to perform the  Clifford operation
\begin{equation}
	\tau \longrightarrow X^{-i}Z^{-j}\ \tau\ Z^{j}X^{i},
\end{equation}
which permutes the smallest component to the origin of the phase space, $W_{\tau}(i,j) \longrightarrow W_{\tau}(0,0)$.
This operation would inevitably affect the magic distillation protocol, but since we consider the state distillation modulo Clifford operations, we are allowed to perform the reverse Clifford operation keeping the magic distillation process as is. \nick{statement here is inaccurate}

The rescaled distribution components are in general given by
\begin{equation}
	W_{\rho_n|\tau_n} = \left(\frac{-v}{W_{\tau}(0,0)}\right)^{i_{\bmo}} \prod_{\bmx \neq \bmo} \left(\frac{u}{W_{\tau}(\bmx)}\right)^{i_{\bmx}},
\end{equation}
where the (integer) indices obey the following conditions:
\begin{align}
&0 \leq i_{\bmx} \leq n \mbox{ for all } \bmx \in \mathbb{Z}_3^2, \nonumber \\
&\sum_{\bmx \in \mathbb{Z}_3^2} i_{\bmx} = n.
\end{align}
It is now easy to find the largest rescaled component.
Firstly, note that $n$ is even, so we require that $i_{\bmo} \in \{0,2,\dots,n\}$ for the component to be positive.
Then, we have that $v \geq u$ because $\epsilon \leq 3/7$, and we have already ensured that $W_{\tau}(0,0) \leq W_{\tau}(\bmx)$ for all $\bmx \in \mathbb{Z}_3^2$.
Therefore, the largest rescaled component occurs when $i_{\bmo} = n$ and $i_{\bmx} = 0$ for $\bmx \neq \bmo$ and is equal to $(v/W_{\tau}(0,0))^n$.
Accordingly, the coordinates of the first Lorenz curve point after the origin are given by
\begin{equation}
	(x_0, L_0) = ((W_{\tau}(0,0))^n, v^n)
\end{equation}

Given a Hamiltonian decomposition,
\begin{equation}
	H = \sum_{k \in \mathbb{Z}_3} E_k\ketbra{E_k}{E_k},
\end{equation}
we can re-express the coordinate location, by writing
\begin{align}
	W_{\tau}(0,0) &= \frac{1}{3\Z}\tr\left[ A_{0,0}e^{\beta H} \right] \nonumber\\
	&= \frac{e^{\beta F}}{3} \tr\left[ \sum_{r \in \mathbb{Z}_3} \ketbra{-r}{r}\sum_{k \in \mathbb{Z}_3} e^{\beta E_k}\ketbra{E_k}{E_k} \right] \\
	&= \frac{e^{\beta F}}{3} \sum_{r \in \mathbb{Z}_3} \alpha_k e^{\beta E_k}
	= \frac{e^{\beta F}}{3} \zeta \\
	&= \frac{e^{\beta (F - \phi)}}{3},
\end{align}
where $\alpha_k, \zeta$ and $\phi$ are as defined prior to the statement of the theorem.
Finally, the coordinates can be expressed as
\begin{equation}
	(x_0, L_0) = \left( \frac{e^{n\beta (F - \phi)}}{3^n}, v(\epsilon)^n \right).
\end{equation}
In the same manner, we can derive the first point coordinates of the output Lorenz curve as
\begin{equation}
	(x'_0, L'_0) = \left( \frac{e^{n\beta (F' - \phi')}}{3^n}, v(\epsilon')^n \right).
\end{equation}

If the largest rescaled component of a state is distinct with no multiplicities, then these coordinates correspond to the first elbow of the corresponding Lorenz curve, whereas if it appears multiple times, then the coordinates derived correspond to a point on the interior of the line segment connecting the origin to the first elbow.
In both cases, the distillation bound remains the same, as is clear by its derivation in~\cref{app:elb_constraints}, and it is given by $L_0 / x_0 \geq L'_0 / x'_0$, which can be directly rearranged into the stated bound of the theorem.
\end{proof}

### 2021-05-04 magic_paper.tex

Within the Wigner representation, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:campbell}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed (free) quantum operations, i.e. the ones that are considered free. The assumption we require on free operations is simply that they send any free state to another free state, meaning that a resource state cannot be created freely, a common assumption among all magic theories.

@@

The mana bound can be directly calculated as
\begin{equation}
	R \leq \frac{\mana{\rho_{\rm{S}}(\epsilon)}}{\mana{\rho_{\rm{S}}(0)}} = 1 + \frac{\ln \left(1 - \frac{8}{15}\epsilon \right)}{\ln \frac{5}{3}}.
\end{equation}
For the noisy Strange state, the max--thauma bound coincides with the mana bound, and are both looser than the majorization bound we derived via the first elbow constraint as illustrated in~\cref{fig:distill_bounds}. 
Numerical, ``optimal'', bounds $R_{\rm{opt}}$ obtained by considering the entirety of the Lorenz curves have been included in the plot as well, highlighting that the first elbow constraint does not capture the full capabilities of majorization.
Notably, in contrast to the other bounds, $R_{\rm{opt}}$ depends explicitly on the copies $m,n$, rather than their ratio $m/n$, so it is specific to the process.
An interesting question arises regarding the value of $R_{\rm{opt}}$ as $n,m \rightarrow \infty$, while $m/n$ remains constant.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.45]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation bounds in the unital fragment.} Distillation bounds obtained by the first elbow majorisation constraint, mana and max--thauma are plotted for $\epsilon' = 0$, up to noise level $\epsilon = 3/7$.
    Majorisation provides stricter rates than the global mana and max--thauma bounds.
    Numerical bounds $R_{\rm{opt}}$ obtained by full Lorenz curve comparison for some Strange state purification processes with even number of copies $m,n \leq 30$ highlight that the first elbow constraint can be improved.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\newpage
\section{Free energy dependent bounds for magic state distillation protocols}
\label{sec:stab}

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \ddd{[cite: look up papers by Marcus Mueller. He has about 3 or 4 on this. Also Mark Wilde has a paper on it]}

@@

This provides a simple means to specify energetic details of a protocol through how much it would disturb some reference equilibrium state. The approach, when combined with relative majorization, also allows us to derive magic distillation bounds that combine both the computational measures $\epsilon,\epsilon'$, with terms that depend on the Hamiltonian of the physical system. As discussed in section (blah), from the perspective of majorization the presence of magic can be viewed as a form of non-classical free energy and so it is also of interest to make an explicit link between the amount of magic in a non-equilibrium state $\rho_S(\epsilon)$ and the actual free energy of an equilibrium state $\tau(\beta)$.

@@

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol on qutrits that transforms $n$ copies of an $\epsilon$--noisy Strange state into $m$ copies of an $\epsilon'$--noisy Strange state, with depolarising errors $\epsilon' \leq \epsilon \leq 3/7$.

	Let $T =(k\beta)^{-1}$ be any temperature for the physical system and let $H= \sum_{k \in \mathbb{Z}_3} E_k |E_k\>\<E_k|$ be the Hamiltonian of each qutrit subsystem in its eigen-decomposition.
Assume that for $n,m \rightarrow \infty$ the protocol's channel generates negligible correlations on the equilibrium state $\tau$, and so $\tau^{\otimes n} \longrightarrow \tau^{\prime \otimes m}$ for $n,m \gg 1$. We write the state $\tau'$ as $\tau' = e^{-\beta H'}/\Z'$ for some Hermitian $H'$.

Given this, there are local Clifford changes of basis $C_1,C_2$, such that $\rho_1 := C_1 \rho_S(\epsilon) C_1^\dagger$ and $\rho_2 := C_2 \rho_S(\epsilon') C_2^\dagger$ for which the protocol gives
\begin{equation}
\rho_1^{\otimes n} \longrightarrow \rho_2^{ \otimes m}.
\end{equation}
Moreover, the magic distillation rate $R = m/n$ for the protocol is bounded by the expression
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\phi' - F')},
\end{equation}
where $F$ is the free energy of $\tau$,  and 
\begin{equation}
\phi = -\beta^{-1} \log \zeta
\end{equation}
with $\zeta$ given by the equations
\begin{align}
\zeta &= \sum_{k\in \mathbb{Z}_3} \alpha_k e^{-\beta E_k}, \nonumber\\
\alpha_k &= \sum_{r \in \mathbb{Z}_3} \braket{E_k}{-r}\braket{r}{E_k}.
\end{align}
The primed variables are defined similarly for the output system.
\end{theorem}
The proof of this is provided in Appendix \ref{free-energy-bound-proof}, and is a generalization of the analysis for unital protocols.

@@

The above analysis makes simplifying assumptions that could easily be dropped, at the price of more complex expressions. For example, we can obtain general expressions without any assumptions on $\epsilon,\epsilon'$ or drop the Clifford change in basis. 
We could also perform similar analysis for general qudits, and different choices of magic states. It might also be of interest consider other choices of reference states that may include more appropriate hardware physics, for example for photonic set-ups.

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality \ddd{[cite]} for approximating the free energy of a system, to obtain bounds in a similar form.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}
\subsection{Variation of bounds with physical parameters}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

@@

\ddd{[Need some numerics on varying the $F'$ case]}
\nick{FIX} The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.

@@

\ddd{[Initial text. To be edited.]}
We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Our bounds only exploited the very simplest aspects of the Lorenz curves of the quasi-distributions, and so it would be of interest to sharpen these bounds and exploit more of the structure in the asymptotic limit. This raises interesting questions that do no arise for probability distributions, because we no longer have a notion of typicality and the central limit theorem does not apply. That said, for special states such as the Strange state the asymptotic behaviour is relatively simple and so one could study this and obtain a better handle on the $n\rightarrow \infty$ behaviour in a way. In appendix (\ddd{BLAH} we give some basic relations on this that could be explored further. \ddd{[where did the remainder term bounds go? People could potentially build on this]}

One might wonder if this approach could also provide \emph{lower bounds} for distillation protocols. Now, in contrast to the upper bounds, it is necessary to include the symplectic constraints on the phase space into the majorization relations.

  For example, the Clifford group action on a quantum system corresponds to the action of the affine symplectic group \ddd{[cite Gross discrete Hudson paper, and some maths source on symplectic group]} on the discrete phase space. Therefore, any convex mixture of Clifford unitaries will correpond to a convex mixture of these group actions. From the Hardy-Littlewood theorem \ddd{[cite]}, we know that traditional majorization is obtained from convex mixtures of arbitrary permutations. However this has been generalised to convex mixtures of elements of a group $G$ that acts on vectors. This $G$--majorization has been studied in the classical literature and a range of results are known about it \ddd{[cite papers]}. For example, if $G$ is a finite reflection group then the $G$--majorization pre-order is described by a finite list of conditions\ddd{[cite]}, just as happens for the full majorization ordering. One route is therefore to consider reflection sub-groups of the affine symplectic group, for which the majorization relation reduces to a finite set of conditions. This would therefore provide concrete lower bounds on distillation rates.

  The topic of $G$--majorization has been extensively studied, but to our knowledge there hasn't been work on relative $G$--majorization. This would correspond to transformations that are \emph{not} unital, but still respect the symplectic structure. Physically, this regime would correspond to a form of thermo-majorization obtained from looking at the action of the Clifford group at a micro-canonical level, and then reducing to a small subsystem \ddd{[cite stat-mech book, cite Matteo review]}. While this seems like a painful thing to consider, there is motivation for this beyond the aim of constructing lower bounds for magic protocols: in the case of statistical mechanics on a phase space this is precisely the situation, albeit in the continuum limit. Statistical mechanics of actual systems obey Hamiltonian dynamics and so automatically respect a symplectic form \ddd{[cite a stat-mech book]}. Therefore the pre-order of statistical mechanical states with respect to phase space dynamics preserving a Gibbs state must correspond to a symplectic thermo-majorization condition. Of course, technical features arise in the continuum limit when considering distributions on an unbounded phase space, but this could be remedied by either considering a compact phase space (e.g. for a particle on a ring) or by first studying the discrete case. Such scenarios also arise in the Quantum Hall Effect \ddd{[cite]} and so this is another regime in which these techniques might be of use.

Finally, it would be of interest to consider the possibility of applying these techniques to other scenarios in which one wishes to distinguish classical from non-classical behaviour based on quasi-probability representations \ddd{[cite quantum optics reference, cite Ferrie review on quasi-reps]}

>>>>

Within the Wigner representation, all negatively represented states are magic states, although in general the converse is not true~\cite{cit:campbell}. However, it is known that all positively represented states used in Clifford circuits admit an efficient classical description, so it is natural to focus on the negatively represented magic states. The free states in any magic theory must therefore lie in the set
\begin{equation}
    \Fmax \coloneqq \{ \rho: \W[\bmx]{\rho} \geq 0 \text{ for all } \bmx \in \cal{P}_d\}.
\end{equation}
Our focus is on states with negativity, so the particular choice of free states is not critical for our analysis. The remaining component that defines any magic theory is the set of allowed quantum operations, i.e. the ones that are considered free. The assumption we require on free operations is simply that they send any free state to another free state, meaning that a resource state cannot be created freely, a common assumption among all magic theories.

@@

The mana bound can be directly calculated as
\begin{equation}
	R \leq \frac{\mana{\rho_{\rm{S}}(\epsilon)}}{\mana{\rho_{\rm{S}}(0)}} = 1 + \frac{\ln \left(1 - \frac{8}{15}\epsilon \right)}{\ln \frac{5}{3}}.
\end{equation}
For the noisy Strange state, the max--thauma bound coincides with the mana bound, and are both looser than the majorization bound we derived via the first elbow constraint as illustrated in~\cref{fig:distill_bounds}. 
Numerical, ``optimal'', bounds $R_{\rm{opt}}$ obtained by considering the entirety of the Lorenz curves have been included in the plot as well, highlighting that the first elbow constraint does not capture the full capabilities of majorization.
Notably, in contrast to the other bounds, $R_{\rm{opt}}$ depends explicitly on the copies $m,n$, rather than their ratio $m/n$, so it is specific to the process.
An interesting question arises regarding the value of $R_{\rm{opt}}$ as $n,m \rightarrow \infty$, while $m/n$ remains constant.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.45]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation bounds in the unital fragment.} Distillation bounds obtained by the first elbow majorisation constraint $R$, mana $\M$ and max--thauma $\theta_{\rm{max}}$ are plotted for $\epsilon' = 0$, up to noise level $\epsilon = 3/7$.
    Majorisation provides stricter rates than the global mana and max--thauma bounds.
    Numerical bounds $R_{\rm{opt}}$ obtained by full Lorenz curve comparison for some Strange state purification processes with even number of copies $m,n \leq 30$ highlight that the first elbow constraint can be improved.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

\section{Free energy dependent bounds for magic state distillation protocols}
\label{sec:stab}

@@

The case of correlations between output subsystems can also be considered, but would lead to a more complex analysis within our majorization framework of Wigner distributions. In this direction, we highlight recent work in majorization in which stochastic independence and correlations are analysed. In particular, it has been shown that stochastic independence (no correlations) of independent distributions can be viewed as a resource in an extension of catalytic majorization, and leads to a single-shot operational interpretation of the Shannon entropy. \nick{cite David?}

@@

This provides a simple means to specify energetic details of a protocol through how much it would disturb some reference equilibrium state. The approach, when combined with relative majorization, also allows us to derive magic distillation bounds that combine both the computational measures $\epsilon,\epsilon'$, with terms that depend on the Hamiltonian of the physical system. As discussed in section~\nick{REF}, from the perspective of majorization the presence of magic can be viewed as a form of non-classical free energy and so it is also of interest to make an explicit link between the amount of magic in a non-equilibrium state $\rho_S(\epsilon)$ and the actual free energy of an equilibrium state $\tau(\beta)$.

@@

\begin{theorem}\label{thm:free-energy}
	Consider a magic distillation protocol on qutrits that transforms $n$ copies of an $\epsilon$--noisy Strange state into $m$ copies of an $\epsilon'$--noisy Strange state, with depolarising errors $\epsilon' \leq \epsilon \leq 3/7$. Let $T =(k\beta)^{-1}$ be any temperature for the physical system and let $H= \sum_{k \in \mathbb{Z}_3} E_k |E_k\>\<E_k|$ be the Hamiltonian of each qutrit subsystem.

Assume for $n,m \rightarrow \infty$ that the protocol's channel $\E$ generates negligible correlations on the equilibrium state, so there is a state $\tau'$ such that
\begin{equation}
	\tau^{\otimes n} \longrightarrow \E(\tau^{\otimes n}) = \tau^{\prime \otimes m}
\end{equation} 
for $n,m \gg 1$. 
We write the output state as $\tau' = e^{-\beta H'}/\Z'$ for some Hermitian $H'$.

Given this, there are local Clifford changes of basis $C_1,C_2$, such that $\rho_1 := C_1 \rho_S(\epsilon) C_1^\dagger$ and $\rho_2 := C_2 \rho_S(\epsilon') C_2^\dagger$ for which the protocol gives
\begin{equation}
\rho_1^{\otimes n} \longrightarrow \rho_2^{ \otimes m}.
\end{equation}
Moreover, the magic distillation rate $R = m/n$ for the protocol is bounded by the expression
\begin{equation}\label{eq:rate_bounds_proof}
	R \leq \dfrac{\ln{\big( 1-\frac{4}{3}\epsilon \big)} + \beta (\phi - F)}{\ln{\big( 1-\frac{4}{3}\epsilon' \big)} + \beta (\phi' - F')},
\end{equation}
where $F$ is the free energy of $\tau$,  and 
\begin{equation}
\phi = -\beta^{-1} \log \zeta
\end{equation}
with $\zeta$ given by the equations
\begin{align}
	\zeta &= \sum_{k\in \mathbb{Z}_3} \alpha_k e^{-\beta E_k}, \nonumber\\
\alpha_k &= \sum_{r \in \mathbb{Z}_3} \braket{E_k}{-r}\braket{r}{E_k}.
\end{align}
The primed variables are defined similarly for the output system.
\end{theorem}
The proof of this is provided in~\cref{free-energy-bound-proof}, and is a generalization of the analysis for unital protocols.

@@

The above analysis makes simplifying assumptions that could easily be dropped, at the price of more complex expressions. For example, we can obtain general expressions without any assumptions on $\epsilon,\epsilon'$ or drop the Clifford change in basis. 
We could also perform similar analysis for general qudits, and different choices of magic states. It might also be of interest to consider other choices of reference states that may include more appropriate hardware physics, for example for photonic set-ups.

The only non-trivial assumption is the assumption that we can ignore correlations in the thermodynamic limit. This is a common situation in thermodynamics, where correlations are often sub-linear in the particle number in the macroscopic limit. One could perturb about this scenario, and make use of variational tools such as the Bogoliubov inequality~\cite{bogolyubov_1966} for approximating the free energy of a system, to obtain bounds in a similar form.

\subsection{Variation of bounds with physical parameters and constraints from complete positivity}
The bounds derived are based on analysis of only part of the Lorenz curves and can be improved via a finer analysis.
This is apparent by simple numerical calculations on the entirety of the curves as seen in~\cref{fig:distill_bounds} for the case of unital operations.
Specifically, the existing bounds follow by considering the dominant terms in the rescaled Wigner distribution and do not, for example, take into account the Lorenz curve's peak structure.

\ddd{[To be tidied+blended]} We also note that one can obtain another set of general distillation bounds from the monotonicity of discrimination measures under quantum channels. For example, the $L_1$ norm is contractive under quantum channels so that 
\begin{equation}
|| \rho - \sigma||_1 \ge ||\E(\rho) - \E(\sigma)||_1,
\end{equation}
for any quantum channel $\E$. Therefore, beyond the constraint that the magic protocol corresponds to a stochastic map in the Wigner distribution, we also have constraints coming from complete-positivity and trace-preservation. For example, if $D(\rho,\sigma)$ is a contractive measure on pairs of quantum states that is also additive, in the sense that $D(\rho^{\otimes n}, \sigma^{\otimes n}) = n D(\rho, \sigma)$, then it is readily seen that within the context of the above theorem we also have,
\begin{equation}
R \le \frac{D(\rho_S(\epsilon),\tau)}{D(\rho_S(\epsilon'), \tau')},
\end{equation}
provided the denominator does not vanish. In particular we can choose the Renyi-divergences $D_\alpha(\rho ||\tau)$....or more concretely the von Neumann divergence $S(\rho ||\tau) = \tr[ \rho \log \rho - \rho \log \tau]$. However, if $\tau$ is taken to be a Gibbs state then $S(\rho ||\tau) = \beta (F[\rho] - F[\tau])$, namely the difference in free energies of the two states. Here we define
\begin{equation}
F[\rho] := \tr[H \rho] - \beta^{-1} S(\rho).
\end{equation}
  Therefore, simply from the fact that the protocol defines a channel, we obtain the bound
\begin{equation}
R \le \frac{F[\rho_S(\epsilon)] - F[\tau]}{F[\rho_S(\epsilon')] - F[\tau']},
\end{equation}
and so we can compare this with the above bounds \ddd{[Some numerics on this to compare?]}

@@

\nick{FIX} The above bounds mainly demonstrate how one can use relative majorization of quasi-distributions to incorporate additional information into distillation bounds. However to unpack the results more it is useful to consider the case in which $H=H'$, and look at $\epsilon' \rightarrow 0$. For this we obtain~\cref{fig:rate_contour} which gives a contour plot of $R$ as a function of inverse temperature $\beta$ and depolarising noise level $\epsilon$ for the initial noisy magic state.
\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.4]{figs/rate_scatter.pdf}
    \caption{\textbf{Temperature-dependent bounds for magic distillation.}
  Shown is a contour plot of  $R(\epsilon, \beta)$ for $H=H'$, where $\beta$ is an inverse temperature and $\epsilon$ is the depolarising noise of the magic state. The relevant data from the spectrum of the qutrit operator $H$ enters in via $\epsilon_\star(\beta)$ and $\beta_\star$.  The vertical dashed line is the `Landauer-like' temperature threshold $\beta_\star$ and the diagonal dashed curve corresponds to the noise threshold $\epsilon_\star (\beta)$ for $\beta \leq \beta_\star$. The $\beta =0 $ line corresponds to the regime in which the magic protocol gives a unital channel.
    }
    \label{fig:rate_contour}
\end{figure}

@@

\ddd{[Initial text. To be edited.]}
We have described how relative majorization can be used to establish upper bounds on magic distillation protocols that take into account additional physics of the system. Given this, one might wonder if this approach could also provide \emph{lower bounds} for distillation protocols. Now, in contrast to the upper bounds, it is necessary to include the symplectic constraints on the phase space into the majorization relations.

  For example, the Clifford group action on a quantum system corresponds to the action of the affine symplectic group on the discrete phase space. Therefore, any convex mixture of Clifford unitaries will correpond to a convex mixture of these group actions. From the Hardy-Littlewood theorem, we know that traditional majorization is obtained from convex mixtures of arbitrary permutations. However this has been generalised to convex mixtures of elements of a group $G$ that acts on vectors. This $G$--majorization has been studied in the classical literature and a range of results are known about it. For example, if $G$ is a finite reflection group then the $G$--majorization pre-order is described by a finite list of conditions, just as happens for the full majorization ordering. One route is therefore to consider reflection sub-groups of the affine symplectic group, for which the majorization relation reduces to a finite set of conditions. This would therefore provide concrete lower bounds on distillation rates.

  The topic of $G$--majorization has been extensively studied, but to our knowledge there hasn't been work on relative $G$--majorization. This would correspond to transformations that are \emph{not} unital, but still respect the symplectic structure. Physically, this regime would correspond to a form of thermo-majorization obtained from looking at the action of the Clifford group at a micro-canonical level, and then reducing to a small subsystem. While this seems like a painful thing to consider, there is motivation for this beyond the aim of constructing lower bounds for magic protocols: in the case of statistical mechanics on a phase space this is precisely the situation, albeit in the continuum limit. Statistical mechanics of actual systems obey Hamiltonian dynamics and so automatically respect a symplectic form. Therefore the pre-order of statistical mechanical states with respect to phase space dynamics preserving a Gibbs state must correspond to a symplectic thermo-majorization condition. Of course, technical features arise in the continuum limit when considering distributions on an unbounded phase space, but this could be remedied by either considering a compact phase space (e.g. for a particle on a ring) or by first studying the discrete case.

In this work we have looked at magic through the lens of thermodynamics. Specifically, we find that it can be viewed as a strictly non-classical form of non-equilibrium free energy.

Our work also raises a fundamental question that is also of interest in statistical mechanics. Specifically: what are the majorization conditions that respect the symplectic structure of phase space?
